---
title: Beer Market Analysis
subtitle: "Data-Driven Mastery: Unlocking Business Potential"
author: 
  - Brett Patmore

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(dplyr)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Beer Market Data
```{r}
beer_markets <- read_csv("https://bcdanl.github.io/data/beer_markets_all.csv")
```


```{r}
#| echo: false
#| eval: true
rmarkdown::paged_table(beer_markets)
```


<br>


## Variable Description

- 'hh': an identifier of the household
  
- 'X_purchase_desc': details on the purchased item
  
- 'quantity': the number of items purchased
  
- 'brand': Bud Light, Busch Light, Coors Light, Miller Lite, or Natural Light
  
- 'dollar_spent': total dollar value of purchase
  
- 'beer_floz': total volume of beer, in fluid ounces
  
- 'price_per_floz': price per fl.oz. (i.e., beer spent/beer floz)

- 'container': the type of container

- 'promo': Whether the item was promoted (coupon or otherwise)

- 'market': Scan-track market (or state if rural)

- 'state': US State

- demographic data, including gender, marital status, household income, class of work, race, education, age, the size of household, and whether or not the household has a microwave or dishwasher

<br>


# Introduction

## Background: Provide context for the research topics, explaining why they are significant or relevant. 

## Statement of the Problem: Clearly articulate the specific problem or issue the project will address.


<br><br>


# Exploratory Data Analysis

## List the questions you aim to answer.

## Address the questions through summary statistics, data visualization, and data transformation.

## Question 1

- Question 1

*Answer:* 
```{r}

  
```


```{r}

```


```{r}

```


<br><br>


# Significance of the Project

## Explain its implications for real-world applications, business strategies, or public policy. 


<br><br>


# References

## List all sources cited in the project. 
